---
title: "Lessons from R4DS"
author: "Doug Duncan"
date: "October 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install pre-reqs

#install.packages(c("nycflights13", "gapminder", "Lahmann"))
pkgs <- c("tidyverse", "nycflights13", "gapminder", "Lahman")

lapply(pkgs, require, character.only = T)

```

# Part I - EXPLORE  
The goal of this part is to understand *data exploration*  

1. Visualization  

2. Transformation - Key Verbs to select, filter, create, and summarize  

3. Answer questions using **transformation** and **visualization**  

**Note** Answers to exercises are [here](https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html)  

## Chap. 1 - Data Visualization with ggplot2  

### First Steps  

**Question to Answer** - do cars with big engines use more fuel than cars with small engines?  

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

```

The *mapping* argument is always paired with aes(), and the args for aes specify the x and y values.

### Exercises  
1. Q: Run ggplot(data = mpg).  What do you see?  **Answer** - an empty plot

```{r}
ggplot(data = mpg)
```

      
2. How many rows are in mtcars?  How many columns?  **Answer** 32 rows, 11 columns

```{r}
mtcars
```


3. What does the drv variable describe?  The type of drivetrain - front, rear, 4wd

```{r}
# ?mpg
```

4. Make a scatterplot of hwy vs cyl

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```

5. What happens if you make a scatterplot of class vs drv?  Why is the plot not useful?  **Answer** Because this creates a scatterplot from 2 categorical variables.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```

### Aesthetic Mappings  

An **aesthetic** is a visual property of the objects in your plot.  Things like size, shape, or the color of your points.  

This plot maps the colors of the points in the **class** variable to reveal the class of each car.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

This plot maps the **size** discrete variable to the class (not recommended). 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```


Here, we map class to the alpha aesthetic, which controls the transparency of the points, or their shape.  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```


### Exercise 1.2  

1. What is wrong with the code below?  **Answer** A parenthese should be after hwy and removed from the end.  Color goes outside of aes!!


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

```


2. Which variables in mpg are categorical?  Which are continuous?  Categorical = (manu, model, year, cyl, trans, drv, fl, class)

```{r}
str(mpg)
```


3. Map a continuous variable to color, size, & shape.  How are these different for categorical vs continuous variables?  **Answer** For continuous variables, there is one color with different shades of the same color.  For categorical variables, there are different colors for each categorical variable.


```{r cars}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = displ, color = cty))


ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = displ, color = fl))
```

4. What happens if you map the same variable to multiple aesthetics?  **Answer** Plot works, but does not provide any additional information.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = displ, color = hwy))
```

5. What does the *stroke* aesthetic do?  What shapes does it work with? **Answer** Modifies the width of the border of shapes.  

```{r}
ggplot(data = mpg, aes(hwy, displ)) + 
  geom_point(shape = 21, colour = "black", fill = "yellow", size = 5, stroke = 2)
```


6. What happens if you map an aesthetic to something other than a variable name like ase(color = displ < 5)?
**Answer** Cool!  Shows values less than 5 in one color, >= 5 in another!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = displ, color = displ < 5))
```


## Common Problems  



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
